Master thesis

### 2017-07-18 02-Introduction.Rmd
Bruno 1st review

Tropical forests perturbations, through deforestation and degradation, account for 8.26 billion of carbon dioxyde emissions per year [@Pearson2017]. It represents the third source of greenhouse gas. Besides deforestation has been rightly the focus of worldwide attention, degradation has been less studied and quantified through tropics. Degradation from forest has been estimated to represents 10 time deforestation [@Herold2011] and represents 20% of greenhouse gas emissions in the brazilian Amazon [@Asner2005].

Sutainable forest management in the tropics (i.e. managed selective harvesting of timber) has been widely promoted internationnaly to combat tropical deforestation and degradation [@Zimmerman2012]. Currently tropical logging accounts for one eight of global timber production [@Blaser2011] and is still increasing. Most tropical timber production originates from selective logging, the targeted harvesting of timber from commercial species in a single cutting cycle [@Martin2015].

@@

The question of selective logging impact on tropical forest can be directly related to the emerging field of biodiversity and ecosystem functionning [@Loreau2000]. Tropical forest outstanding biodiversity will be both a factor and a result of forest ecosystem response to logging disturbance. And forest ecosystem response to logging disturbance will directly modify ecosystem functionning in both short and long term. Consequently assessing selective logging effect on tropical forest linking biodiversity and ecosystem seems an obvious and promising way [@Loreau2010].

Negative short term impacts of selective logging have been assessed [@Carreno-Rocabado2012; @DeAvila2015;  but see @Martin2015]. Much less is known about the long term impact [@Osazuwa-Peters2015]. The main reason is the difficulty to conduct long term empirical study [but see @Herault2010], which can be completed by the use of forest simulators [@Huth2004; @Khler2004; @Ruger2008; @Tietjen2006]. Individual-based models of forest dynamics present the perfect framework to develop such joint biodiversity-ecosystem approaches [@Li]. Individual-based models describe forest 'patches' accumulating carbon through time, assessing tree growth within the patch, or releasing carbon through gap opening [@Bugmann2001]. Up to several dozens of different Plant Functional Types (PFTs) are generally defined and models can sometimes be fully spatially explicit [@Pacala1996]. Recently, the forest growth simulator TROLL [@Chave1999], an individual-based and spatially explicit forest model, was developped to introduce recent advances in plant physiological community. TROLL model relates physiological processes to species-specific functional traits [@Li]. Consequently, TROLL model allow to simulate fully a neotropical forest biodiversity to study biodiversity-ecosystem functionning link response to logging disturbance.

>>>>

Tropical forests perturbations, through deforestation and degradation, account for 8.26 billion of carbon dioxyde emissions per year [@Pearson2017]. It represents the third source of greenhouse gas. Besides deforestation has been rightly the focus of worldwide attention, degradation has been less studied and quantified through tropics. Degradation from forest has been estimated to represents 10 time deforestation [@Herold2011] and represents 20% of greenhouse gas emissions in the brazilian Amazon [@Asner2005]. Degradation has met numerous definitions [@Simula2009], but can be defined as the result from a disturbance event that induced a modification of the forest ecosystem reducing ecosystem services while keeping the ecosystem as a forest, contrary to deforestation.

Sutainable forest management in the tropics (i.e. managed selective harvesting of timber) has been widely promoted internationnaly to combat tropical deforestation and degradation [@Zimmerman2012]. Currently logging from tropical forests accounts for one eight of global timber production [@Blaser2011] and is still increasing. Most tropical timber production originates from selective logging, the targeted harvesting of timber from commercial species in a single cutting cycle [@Martin2015].

@@

The question of selective logging impact on tropical forest can be directly related to the emerging field of biodiversity and ecosystem functionning [@Loreau2000]. Tropical forest outstanding biodiversity will be both a factor and a result of forest ecosystem response to logging disturbance. And forest ecosystem response to logging disturbance will directly modify ecosystem functionning in both short and long term. Consequently assessing selective logging effect on tropical forest linking diversity and ecosystem functionning seems an obvious and promising way [@Loreau2010].

Negative short term impacts of selective logging have been assessed [@Carreno-Rocabado2012; @DeAvila2015;  but see @Martin2015]. Much less is known about the long term impact [@Osazuwa-Peters2015]. The main reason is the difficulty to conduct long term empirical study [but see @Herault2010], which can be completed by the use of forest simulators [@Huth2004; @Khler2004; @Ruger2008; @Tietjen2006]. Individual-based models of forest dynamics present the perfect framework to develop such joint biodiversity-ecosystem approaches [@Li]. Individual-based models describe forest accumulating carbon through time, assessing tree growth, or releasing carbon through gap opening [@Bugmann2001]. Up to several dozens of different Plant Functional Types (PFTs) are generally defined and models can sometimes be fully spatially explicit [@Pacala1996]. Recently, the forest growth simulator TROLL [@Chave1999], an individual-based and spatially explicit forest model, was developped to introduce recent advances in plant physiological community. TROLL model relates physiological processes to species-specific functional traits [@Li]. Consequently, TROLL model allow to simulate fully a neotropical forest biodiversity to study biodiversity-ecosystem functionning link response to logging disturbance.

### 2017-07-18 03-Model.Rmd
Bruno 1st review

TROLL model each tree indivdually in a located environment. Thus TROLL model, alongside with SORTIE  [@Pacala1996; @Uriarte2009] and FORMIND [@Fischer2016; @Kohler1998], can be defined as an individual-based and spatially explicit forest growth model. TROLL simulates the life cycle of individual trees from recruitment, with a diameter at breast height (dbh) above 1 cm, to death with growth and seed production. Trees are growing in a located light environment explicitly computed witin voxels of 1 $m^3$. Each tree is consistently defined by its age, diameter at brest height (dbh), height (h), crown radius (CR), crown depth (CD) and leaf area (LA) (see figure \@ref(fig:TROLLtree)). Tree geometry is calculated with allometric equations but leaf area vary dinamically within each crown following carbon allocations. Voxels resolution of 1 $m^3$ allow the establishment of maximum one tree by 1x1 m pixels. Each tree is flagged with a species label inherited from the parent tree through the seedling recruitment. A species label is associated to a number of species specific parameters (see table \@ref(tab:traits)) related to functional trait values which can be sampled on the field.

@@

In french guiana context, sylviculture can be narrow to selective logging, which can be split in two steps: selection and harvesting. Selection encompass choice of the havrestable area, tree designation by the forest office, tree selection by the harvester, and removal off tree probbed as rotten by the lumber. Harvesting encompass tree felling, tracks opening, and long term damages (simplified in gap damages in current TROLL implementation).

>>>>

TROLL model each tree indivdually in a located environment. Thus TROLL model, alongside with SORTIE  [@Pacala1996; @Uriarte2009] and FORMIND [@Fischer2016; @Kohler1998], can be defined as an individual-based and spatially explicit forest growth model. TROLL simulates the life cycle of individual trees from recruitment, with a diameter at breast height (dbh) above 1 cm, to death with growth and seed production. Trees are growing in a spatialized light environment explicitly computed witin voxels of 1 $m^3$. Each tree is consistently defined by its age, diameter at brest height (dbh), height (h), crown radius (CR), crown depth (CD) and leaf area (LA) (see figure \@ref(fig:TROLLtree)). Tree geometry is calculated with allometric equations but leaf area vary dinamically within each crown following carbon allocations. Voxels resolution of 1 $m^3$ allow the establishment of maximum one tree by 1x1 m pixels. Each tree is flagged with a species label inherited from the parent tree through the seedling recruitment. A species label is associated to a number of species specific parameters (see table \@ref(tab:traits)) related to functional trait values which can be sampled on the field.

@@

In french guiana context, sylviculture can be narrow to selective logging, which can be split in two steps: selection and harvesting. Selection encompass choice of the havrestable area, harvestable tree designation by the forest office, harvested tree selection by the harvester, and removal off tree probbed as rotten by the lumber. Harvesting encompass tree felling, tracks opening, and long term damages (simplified in gap damages in current TROLL implementation).

### 2017-07-18 04-MnMs.Rmd
Bruno 1st review

In order to assess the role of biodiversity in ecosystem answer to both disturbance and sylviculture, we needed to create a space of experiments encompassing both variation of disturbance, biodiversity and time. Disturbance was represented by percentage of basal area loss (0%, 25%, 50% and 75%), or as a selective logging simulation using default parameters. Biodiversity was integrated with two of its components: taxonomic and functional diversities. We used species richness $SR$ to represents taxonomic diversity (5, 25, and 125 species). Functional diversity can be related to numerous components, and @Borgy2017 argued for 5: richness, divergence, regularity, overlap and mean. Because mature forest were created from a bare soil with TROLL simulations, we could not control a priori divergence, regularity and overlap but only assess theim before diturbance. Consequently, we focused on functional richness with convex hull volume $CHV$ and functional mean with community weighted mean $CWM$. For each level of species richness $SR$, we selected 20 communities with growing convex hull volume $CHV$ but with a community weighted means close to the regional species pool community weighted means. Effectivelly, we did not wanted drastic change in community means that could have more effect than functional richness itself. This design of experiments resulted in 60 communities ($5~SR*20~CHV$) and 240 simulations ($60 ~communities*4~levels~of~disturbance$) over 600 years (maturity being assumed after 500 years of regeneration [@Li]). Figure \@ref(fig:DOE) presents the design of experiment for communities biodiversity after the mature forest were simulated, and thus before disturbance. We obtained a broad range of both functionl dispersion $FDis$ and aboveground biomass $AGB$ for simulated forest ecosystems before disturbance.

@@

## Ecosystem answer analysis

@@

Tropical forest ecosystems provides numerous ecosystem services linked to several ecosystem functions. We decided to describe simiulated tropical forests in two major functions: forest dynamic and forest production. Forest dynamic was represented by aboveground biomass ($AGB$ in $ton~C.ha^{-1}$), basal area ($BA$ in $m^2.ha^{-1}$), total number of stem ($N$), number of stem above $10~cm$ diameter ($N10$), and number of stem above $30~cm$ diameter ($N30$). Forest production was represented by growth pimary productivity ($GPP$ in $MgC.ha^{-1}$), net pimary productivity ($NPP$ in $MgC.ha^{-1}$), tree autotrophic respiration in day ($Rday$ in $MgC.ha^{-1}$) and tree autotrophic respiration in night ($Rnight$ in $MgC.ha^{-1}$).

@@

The resilience of the system $R(t)$ at the time $t$ is described by the ratio of recovery $Recovery(t)$ at time $t$ to loss suffered $Loss(t_d)$ at disturbance time $t_d$. But in our peculiar case of tropical forest ecosystems, the equilibrium used to calculate $Loss(t_d)$ can not be reduced to a specific time if the equilibrium is dynamic. Consequently, to encompass undisturbed ecosystem variations throught time, we simulated an undisturbed control ecosystem $C$. And the resilience of the system $R(t)$ at the time $t$ was defined as the ratio of the ecosystem metric values in the disturbed simulation $X_T(t)$ over the ecosystem metric values from the control $X_C(t)$ . Thus, the value of resilience $R(t)$ is normalized for all simulations and metrics. $R(t)$  will be equal to $R_{eq} = 1$ when reaching the equilibrium value. Consequently we can calculate an euclidean distance to equilibrium $d_{eq}(t)$ as $d_{eq}(t) = \sqrt{(R_{eq} - R(t))^2}$. Ecosystem euclidean distance to equilibrium was calculated in a multi-dimensional space for the two functions described above: forest dynamic (AGB, BA, N, N10, and N30) and forest production (GPP, NPP, Rday, and Rnight). We then used cummulative integral over time of eulcidean distance to equilibrium to assess simulations resilience.

@@

$X_{sp}$ is the ecosystem variable value for one species either in mixture $X_{sp}(mixture)$ or in monoculture $X_{sp}(monoculture)$. $P_{sp}$ is the proportion of the species in the mixture represented by species relative abundance. Consequently, $CE$ averages diversity effects of all species presents in the mixture (both negatives and positives). Whereas $SE$ become positive when dominant species outperform theimselves in mixture than in monoculture, and negative when less dominant species outperform theimselves in mixture than in monoculture [@Tobner2016]. But similarly to resilience measurement, biodiversity net effect $NE$ can be in a dynamic equilibrium and vary over time without disturbance. So in order to correctly assess selection and complementarity effect in answer to disturbance, we normalized it by undisturbed control ecosystem net effect $NE_C$ to measure treatments net effect resilience $R(NE_T)$:

>>>>

In order to assess the role of biodiversity in ecosystem answer to both disturbance and sylviculture, we needed to create a space of experiments encompassing both variation of disturbance, biodiversity and time. Disturbance was represented by percentage of basal area loss (0%, 25%, 50% and 75%), or as a selective logging simulation using default parameters. Biodiversity was integrated with two of its components: taxonomic and functional diversities. We used species richness $SR$ to represents taxonomic diversity (5, 25, and 125 species). Functional diversity can be related to numerous components, and @Borgy2017 argued for 5: richness, divergence, regularity, overlap and mean. Because mature forest were created from a bare soil with TROLL simulations, we could not control a priori divergence, regularity and overlap but only assess them after running the simulations, i.e. before applying the diturbance. Consequently, we focused on functional richness with convex hull volume $CHV$ and functional mean with community weighted mean $CWM$. For each level of species richness $SR$, we selected 20 communities with growing convex hull volume $CHV$ but with a community weighted means close to the regional species pool community weighted means. Effectivelly, we did not wanted drastic change in community means that could have more effect than functional richness itself. This design of experiments resulted in 60 communities ($5~SR*20~CHV$) and 240 simulations ($60 ~communities*4~levels~of~disturbance$) over 600 years (maturity being assumed after 500 years of regeneration [@Li]). Figure \@ref(fig:DOE) presents the design of experiment for communities biodiversity after the mature forest were simulated, and thus before disturbance. We obtained a broad range of both functionl dispersion $FDis$ and aboveground biomass $AGB$ for simulated forest ecosystems before disturbance.

@@

## Ecosystem response analysis

@@

Tropical forest ecosystems provides numerous ecosystem services linked to several ecosystem functions. We decided to describe simiulated tropical forests in two major functions: forest structure and forest functionning. Forest structure was represented by aboveground biomass ($AGB$ in $ton~C.ha^{-1}$), basal area ($BA$ in $m^2.ha^{-1}$), total number of stem ($N$), number of stem above $10~cm$ diameter ($N10$), and number of stem above $30~cm$ diameter ($N30$). Forest functionning was represented by growth pimary productivity ($GPP$ in $MgC.ha^{-1}$), net pimary productivity ($NPP$ in $MgC.ha^{-1}$), tree autotrophic respiration in day ($Rday$ in $MgC.ha^{-1}$) and tree autotrophic respiration in night ($Rnight$ in $MgC.ha^{-1}$).

@@

The resilience of the system $R(t)$ at the time $t$ is described by the ratio of recovery $Recovery(t)$ at time $t$ to loss suffered $Loss(t_d)$ at disturbance time $t_d$. But in our peculiar case of tropical forest ecosystems, the equilibrium used to calculate $Loss(t_d)$ can not be reduced to a specific time if the equilibrium is dynamic. Consequently, to encompass undisturbed ecosystem variations throught time, we simulated an undisturbed control ecosystem $C$. And the resilience of the system $R(t)$ at the time $t$ was defined as the ratio of the ecosystem metric values in the disturbed simulation $X_T(t)$ over the ecosystem metric values from the control $X_C(t)$ . Thus, the value of resilience $R(t)$ is normalized for all simulations and metrics. $R(t)$  will be equal to $R_{eq} = 1$ when reaching the equilibrium value. Consequently we can calculate an euclidean distance to equilibrium $d_{eq}(t)$ as $d_{eq}(t) = \sqrt{(R_{eq} - R(t))^2}$. Ecosystem euclidean distance to equilibrium was calculated in a multi-dimensional space for the two functions described above: forest strcuture (AGB, BA, N, N10, and N30) and forest functionning (GPP, NPP, Rday, and Rnight). We then used integrated eulcidean distance to equilibrium over time to assess simulations resilience.

@@

$X_{sp}$ is the ecosystem variable value for one species either in mixture $X_{sp}(mixture)$ or in monoculture $X_{sp}(monoculture)$. $P_{sp}$ is the proportion of the species in the mixture represented by species relative abundance. Consequently, $CE$ averages diversity effects of all species presents in the mixture (both negatives and positives). Whereas $SE$ become positive when dominant species outperform theimselves in mixture than in monoculture, and negative when less dominant species outperform theimselves in mixture than in monoculture [@Tobner2016]. But similarly to resilience measurement, biodiversity net effect $NE$ is a dynamic equilibrium and vary over time without disturbance. So in order to correctly assess selection and complementarity effect in answer to disturbance, we normalized it by undisturbed control ecosystem net effect $NE_C$ to measure treatments net effect resilience $R(NE_T)$:

### 2017-07-18 05-Results.Rmd
Bruno 1st review

Most of functional traits had a significant long term influence on ecosystem outputs (figure \@ref(fig:ftVar)). On the other hand, few functional traits influenced final ecosystem structure. Only specific maximum diameter **dmax** add higher diversity for greater orders implying better evenness in species distributions. Regarding functional composition, traits fixed to mean did not change other functional traits density distribution. Moreover, seedrain did not seem to affect aboveground biomass and final ecosystem height and diameter structure. Seedrain constant fixed to 2 or 20 seed per hectare seemed to have a similar effect. Lower seedrain implied faster decrease of stem above 10 cm dbh and higher number of stem above 30 cm diameter at breast height after ecosystem resilience to disturbance (approximately 50 years). Lower seedrain than default decreased basal area over time. In addition, lower seedrain than default decreased equitability by increasing abundance of abundant species and decreasing abundance of less abundant species. See [Appendix 4: Sensitivity analysis] for further details.

```{r ftVar, echo=FALSE, fig.cap='Functional traits effect on simulation ecosystem variations over time. Number of trees with dbh above 10 cm (N10) and 30 cm (N30), above ground biomass (AGB) and basal area (BA). Grey area represents the interval of control replicates whereas black line represents the mean of control replicates.', cache=TRUE, warning=FALSE}
path <- '~/Documents/ECOFOG/Results/sensitivity/control'
load(file.path(path, 'control.Rdata'))
path <- '~/Documents/ECOFOG/Results/sensitivity/traits'
load(file.path(path, 'traits.Rdata'))
n10 <- plot(traits, control, what = 'abu10', ggplot2 = T) +
  scale_color_discrete('Trait fixed\nto mean') +
  ylab('Number of stem above\n10 cm diameter\nat breast height')
n30 <- plot(traits, control, what = 'abu30', ggplot2 = T) +
  scale_color_discrete('Trait fixed\nto mean') +
  ylab('Number of stem above\n30 cm diameter\nat breast height')
agb <- plot(traits, control, what = 'agb', ggplot2 = T) +
  scale_color_discrete('Trait fixed\nto mean') +
  ylab('Aboveground biomass (kgC/ha)')
ba <- plot(traits, control, what = 'ba', ggplot2 = T) +
  scale_color_discrete('Trait fixed\nto mean') +
  ylab('Basal area (m2/ha')
cowplot::plot_grid(n10, n30, agb, ba, labels = c('N10', 'N30', 'AGB', 'BA'), nrow = 2, ncol = 2)
rm(list = ls()) ; invisible(gc())
```

@@

We measured all ecosystem outputs biodiversity net effect in disturbance simulations by comparing theim to their species corresponding monoculture simulations. The net effect was then partition between selection and complementarity effect. We normalized treatments effect by control net effect (see Table \@ref(tab:tNE)) to measure their aboveground biomass resilience to disturbance (see Figure \@ref(fig:gBE)). We found that complementarity effect was recovering net effect in the first decades untill it diminished after a century. On the contrary selection effect was reduced or even removed by the disturbance. It increased during the whole simulation and was greater than complementarity effect only after decades. The time lag for which complementarity effect was greater than selection effect was increasing with disturbance intensity. Finally 600 years after the disturbance event biodiversity net effect was still not recovered for a disturbance intensity greater than 25% of basal area. We obtained those results for aboveground biomass. We found similar results but with an amlified signal for basal area ($BA$) and stem abundance ($N$ but with an inverted signal because of the forest self thinning) (see [Appendix 5: Disturbance simulations]). Finally, forest growth primary productivity ($GPP$) recovered in few years (proportionnaly to disturbance), and its net effect was maintained by complementarity effect (see [Appendix 5: Disturbance simulations]).

>>>>

Most of functional traits had a significant long term influence on ecosystem outputs (figure \@ref(fig:ftVar)). Only specific maximum diameter **dmax** add higher diversity for greater orders implying better evenness in species distributions. Regarding functional composition, traits fixed to mean did not change other functional traits density distribution. Moreover, seedrain did not seem to affect aboveground biomass and final ecosystem height and diameter structure. Seedrain constant fixed to 2 or 20 seed per hectare seemed to have a similar effect. Lower seedrain implied faster decrease of stem above 10 cm dbh and higher number of stem above 30 cm diameter at breast height after ecosystem resilience to disturbance (approximately 50 years). Lower seedrain than default decreased basal area over time. In addition, lower seedrain than default decreased equitability by increasing abundance of abundant species and decreasing abundance of less abundant species. See [Appendix 4: Sensitivity analysis] for further details.

```{r ftVar, echo=FALSE, fig.cap='Functional traits effect on simulation ecosystem variations over time. Number of trees with dbh above 10 cm (N10) and 30 cm (N30), above ground biomass (AGB) and basal area (BA). Sensitivity of model to functional traits was performed  by fixing species trait values to their mean. Grey area represents the interval of control replicates whereas black line represents the mean of control replicates, thus if ecosystems outputs are outside of grey area values the studied parameter is considered to have a significant influence on the model.', cache=TRUE, warning=FALSE}
path <- '~/Documents/ECOFOG/Results/sensitivity/control'
load(file.path(path, 'control.Rdata'))
path <- '~/Documents/ECOFOG/Results/sensitivity/traits'
load(file.path(path, 'traits.Rdata'))
n10 <- plot(traits, control, what = 'abu10', ggplot2 = T) +
  scale_color_discrete('Trait fixed\nto mean') +
  ylab('Number of stem above\n10 cm diameter\nat breast height')
n30 <- plot(traits, control, what = 'abu30', ggplot2 = T) +
  scale_color_discrete('Trait fixed\nto mean') +
  ylab('Number of stem above\n30 cm diameter\nat breast height')
agb <- plot(traits, control, what = 'agb', ggplot2 = T) +
  scale_color_discrete('Trait fixed\nto mean') +
  ylab('Aboveground biomass (kgC/ha)')
ba <- plot(traits, control, what = 'ba', ggplot2 = T) +
  scale_color_discrete('Trait fixed\nto mean') +
  ylab('Basal area (m2/ha')
cowplot::plot_grid(n10, n30, agb, ba, labels = c('N10', 'N30', 'AGB', 'BA'), nrow = 2, ncol = 2)
rm(list = ls()) ; invisible(gc())
```

@@

We measured all ecosystem outputs biodiversity net effect in disturbance simulations by comparing theim to their species corresponding monoculture simulations. The net effect was then partition between selection and complementarity effect. We normalized complementarity and selection effect of disturbed simulations by biodiversity net effect of undisturbed control simulation (see Table \@ref(tab:tNE)) to measure the resilience to disturbance of their aboveground biomass (see Figure \@ref(fig:gBE)). We found that complementarity effect was recovering biodiversity net effect and was stronger than selection effect in the first decades. But after few decades the complementarity effect diminsihed toward a low value. On the contrary selection effect was reduced or even removed by the disturbance. It increased during the whole simulation and was greater than complementarity effect only after decades. The time lag for which complementarity effect was greater than selection effect was increasing with disturbance intensity. Finally 600 years after the disturbance event biodiversity net effect was still not recovered for a disturbance intensity greater than 25% of basal area. We obtained those results for aboveground biomass. We found similar results but with an aplified signal for basal area ($BA$) and stem abundance ($N$ but with an inverted signal because of the forest self thinning) (see [Appendix 5: Disturbance simulations]). Finally, forest growth primary productivity ($GPP$) recovered in few years (proportionnaly to disturbance), and its net effect was maintained by complementarity effect (see [Appendix 5: Disturbance simulations]).

### 2017-07-18 06-Discussion.Rmd
Bruno 1st review

Belowground processes, herbaceaous plants, epyphytes and lianas are not simulated in TROLL but they not reprensent the only limit of the model. Other processes are simulated but simplified, and we used sensitivity analysis to assess their relative importance. We found that few functional traits were influencing whole forest structure and dynamic. In addition, we found that the seed rain constant had an important effect on species functional composition and diversity. High external seed rain resulted in a quick recovery of the system toward an equilibrium close to the regional species frequency levied by the seed rain. On the contrary, low seed rain let the simulated forest works has a closed system with more system feedback but a lower stability through time with longer species diversity transitions. In order to study the role of diversity in forest resilience we decided to remove the seed rain constant to get a closed system and look at the role of diversity when it maintains itself through feedbacks and not with immigration.

@@

Our results validated the hypothesis of a significative relationship between forest resilience and functional diversity and evenness ($p < 0.01$ and $p < 0.05$ respectivelly). Thus we were able to show that diversity improve forest resilience. More particularly, functional diversity seems a major aspect of resilience if its enforced by a high functional evenness. Effectively high functional diversity needs evenness in order to better answer disturbance, if not the diversity is masked by ecosystem dominant species. Those results confirms the review of @Diaz2001 advocating for underevaluated importance of plant functional diversity in ecosystem processes. Additionally, the role of evenness  confirm the review of @Zhang2012 who highlighted the role of evenness in productivity and thus resilience following our hypothesis. Finally, species and functional richnesses is not directly increasing resilience. But increased species and functional richnesses will increase chance for high functional diversity through the sampling effect [@Loreau1998]. An higher sampling of regional species pool will allow a greater chance to pick more functionally diverse species.

@@

We found that complementary effect was insuring forest resilience in the beginning of forest successions. We interpreted this results as the consequence of facilitation processes. As the only ressource simulated by TROLL is the light, the main facilitation will be light shading of post-pioneer species by pioneer species in disturbance gaps. But the complementarity effect is reducing through time, to let the selection effect insure forest resilience due to an inforced dominance of more productive species. The diminishing of complementarity effect is due to competitive selection though time in gaps succession. But the study scale matters as shwon by @Chisholm2013. Here we look at processes to a 16 ha scale. We do not have topography reducing micro-environment effects, but @Chisholm2013 results suggest that complementarity will be stronger at smaller scale and could explain its low value after forest recovery. Finally, our findings confirm results of @Tobner2016 realized on experimental forests of 4 years in low diverse forest of Canada. They also found that selection effect was greater than complementarity in most of the cases. In the case of selective logging the complementarity effect will thus be the major effect between two cutting cycles due to the cycle length (several decades in guyana shield). High forest diversity with important species complementarity through increased functional diversity is thus an advantage for forest recovery between two cutting cycles in order to maintain both productive ecosystem and sustainable management.

@@

We used closed forest system simulated with TROLL forest model to evaluate the role and mechanisms of biodiversity in tropical forest resilience to disturbance. We found that diversity improved forest resilience. Additionally we found that complementarity between species was insuring forest recovery in forest succession start with facilitation, before more productive species dominate the forest and insure forest recovery. Our results advocate for sustainable selective logging against monoculture stand: high diversity will increase forest resilience and thus improve logging cycles. Even if monoculture stand are naturalized they will not reach mature forest diversity, and consequently they will not reach its natural high resilience. But on the other hand selective logging in tropical forest needs to be sustainable, if not the diversity will slowly decrease after each cycles degrading forest ecosystem functions, and resilience will decrease due to negative feedbacks. @Zimmerman2012 criticized the state of sylviculture in tropics, suggesting that "we have not been able to reconcile these opposing biological and economic forces". Still, they advocates for a possibility of sustainable tropical logging, already existing in small-scale, that will need proper funding from the internationnal communtiy. This view correspond to the high resilience of tropical forest hyperdiverse systems, we were able to show, and let hope for a future sustainable selective logging in tropical forests.

>>>>

Belowground processes, herbaceaous plants, epyphytes and lianas are not simulated in TROLL but they not reprensent the only limit of the model. Other processes are simulated but simplified, and we used sensitivity analysis to assess their relative importance. We found that few functional traits were influencing whole forest structure and dynamic. In addition, we found that the seed rain constant had an important effect on species functional composition and diversity. High external seed rain resulted in a quick recovery of the system toward an equilibrium close to the regional species frequency levied by the seed rain. On the contrary, low seed rain let the simulated forest works as a closed system with more system feedback but a lower stability through time with longer species diversity transitions. In order to study the role of diversity in forest resilience we decided to remove the seed rain constant to get a closed system and look at the role of diversity when it maintains itself through feedbacks and not with immigration.

@@

Our results validated the hypothesis of a significative relationship between forest resilience and functional diversity and evenness ($p < 0.01$ and $p < 0.05$ respectivelly). Thus we were able to show that diversity improve forest resilience. More particularly, functional diversity seems a major aspect of resilience if its strengthened by a high functional evenness. Effectively high functional diversity needs evenness in order to better answer disturbance, if not the diversity is masked by ecosystem dominant species. Those results confirms the review of @Diaz2001 advocating for underevaluated importance of plant functional diversity in ecosystem processes. Additionally, the role of evenness  confirm the review of @Zhang2012 who highlighted the role of evenness in productivity and thus resilience following our hypothesis. Finally, species and functional richnesses are not directly increasing resilience. But increased species and functional richnesses will increase chance for high functional diversity through the sampling effect [@Loreau1998]. An higher sampling of regional species pool will allow a greater chance to pick more functionally diverse species.

@@

We found that complementary effect was insuring forest resilience in the beginning of forest successions. We interpreted this results as the consequence of facilitation processes. As the only ressource simulated by TROLL is the light, the main facilitation will be light shading of post-pioneer species by pioneer species in disturbance gaps. We found similar results to @Morin2011 regarding complementarity and light facilitation. But the complementarity effect is reducing through time, to let the selection effect insure forest resilience due to an inforced dominance of more productive species. The diminishing of complementarity effect is due to competitive selection though time in gaps succession. But the study scale matters as shwon by @Chisholm2013. Here we look at processes to a 16 ha scale. We do not have topography reducing micro-environment effects, but @Chisholm2013 results suggest that complementarity will be stronger at smaller scale and could explain its low value after forest recovery. Finally, our findings confirm results of @Tobner2016 realized on experimental forests of 4 years in low diverse forest of Canada. They also found that selection effect was greater than complementarity in most of the cases. In the case of selective logging the complementarity effect will thus be the major effect between two cutting cycles due to the cycle length (several decades in guyana shield). High forest diversity with important species complementarity through increased functional diversity is thus an advantage for forest recovery between two cutting cycles in order to maintain both productive ecosystem and sustainable management.

@@

We used closed forest system simulated with TROLL forest model to evaluate the role and mechanisms of biodiversity in tropical forest resilience to disturbance. We found that diversity improved forest resilience together with productivity [@Liang2016]. Additionally we found that complementarity between species was insuring forest recovery in forest succession start with facilitation, before more productive species dominate the forest and insure forest recovery. Our results advocate for sustainable selective logging against monoculture stand: high diversity will increase forest resilience and thus improve logging cycles. Even if monoculture stand are naturalized they will not reach mature forest diversity, and consequently they will not reach its natural high resilience. But on the other hand selective logging in tropical forest needs to be sustainable, if not the diversity will slowly decrease after each cycles degrading forest ecosystem functions, and resilience will decrease due to negative feedbacks. @Zimmerman2012 criticized the state of sylviculture in tropics, suggesting that "we have not been able to reconcile these opposing biological and economic forces". Still, they advocates for a possibility of sustainable tropical logging, already existing in small-scale, that will need proper funding from the internationnal communtiy. This view correspond to the high resilience of tropical forest hyperdiverse systems, we were able to show, and let hope for a future sustainable selective logging in tropical forests.
